\documentclass[midd]{thesis}

\usepackage{graphicx}
\usepackage{times}
\usepackage{listings}
\usepackage{changepage}
\usepackage{longtable}
\lstset{
    basicstyle=\linespread{1}\ttfamily,
}

\bibliographystyle{plain}

\title {Examining the Security of Local Inter-Process Communication}

\author {Brendan Leech}
\adviser {Professor Peter C. Johnson}

\begin{document}

\maketitle
\pagenumbering{roman}

\begin{abstract}
Local inter-process communication allows processes on the same computer to share data.  Even though this communication stays within a single computer, best practices for computer security must still be followed.  In this thesis, I investigate the security of local inter-process communication for four different applications.  While I did not find bugs in the vast majority of local IPC endpoints that I studied, I did find one bug due to incorrectly parsing input and validating invalid input.  Additionally, I created two Denial-of-Service attacks that prevent a user from opening an application and interacting with much of the Internet.  These three findings underscore the lack of emphasis given to local IPC and highlight the need to focus more on its security in the future.
\end{abstract}

\begin{acknowledgements}
I'd like to thank my advisor, Pete, for all of his help this semester and throughout the six other classes I have been lucky enought to take with him.  This project would not have been a success without his advice and guidance throughout the past four years and particularly throughout this thesis process.  I also could not have made it this far in my computer science education without the other amazing professors I have had here at Middlebury College, especially Professor Briggs and Professor Christman, each of whom I took multiple classes with throughout my time in the Middlebury CS department.  I'd also like to thank the people who took my survey, which helped inform my decision of what applications to investigate.  Finally, I want to thank my friends and family who supported me throughout the semester, especially those who helped run some of my tests as time was running out.
\end{acknowledgements}

\contentspage
\tablelistpage   % comment this out if you don't have any tables
\figurelistpage

\normalspacing \setcounter{page}{1} \pagenumbering{arabic}

\input{introduction}
\input{ipc}
\input{ipcSecurity}
\input{methods}
\input{results}
\input{conclusion}
\input{discussion}

\appendix
\input{appendixCode}
\input{appendixGroups}
\input{appendixDataTable}


\bibliography{thesis}

\end{document}
